Can Auxiliary Indicators Improve COVID-19 Forecasts and Hotspot Prediction?

### 2021-06-12 forecast/paper/paper.tex
Edits up through methods A

\title{Can COVID-19 Indicators Improve Forecasts and Hotspot Prediction?}

@@

The indicators we focus on provide information not generally available from
standard public health reporting. The COVID Tracking and Impact Survey (CTIS),
run by the Delphi group in partnership with Facebook, reaches roughly 50,000
respondents across the United States daily, and provides information on
symptoms, social distancing, and other attitudes and behaviors related to
COVID. The particular indicator we examine, which we refer to as \fb, is the
estimated percentage of people who know someone with COVID-like illness in their
the local community.

In some ways, this can be seen as a
proxy for the strength of social contacts between infected and susceptible
populations, a key feature for disease modeling~\cite{Zhang1481} The Google
indicator, which we refer to as \gs, tracks the local intensity of Google searches related to lack-of-smell
(anosmia) or lack-of-taste (ageusia), both of which have been connected to
COVID19~\cite{KlopfenAnosmia,vaira2020anosmia}. Two
indicators are derived from de-identified insurance claims data from Change Healthcare, a large
healthcare technology company that aggregates data from many providers. The
first signal, which we refer to as \chngcov, is the estimated percentage of outpatient doctor visits with
confirmed COVID-19 diagnosis (potentially collected earlier than the related
report to the local health authority), while the second, which we
refer to as \chngcli, is the estimated percentage of outpatient
visits with COVID-like illness. Finally, \attn{describe \dv~indicator, perhaps best described in conjunction with \chngcov}.  The hope is that each
of these signals serves as a ``leading indicator'' of future COVID cases,
improving the forecastibility of COVID cases and hotspots. For more details on
these and other indicators, see the companion paper.

% RJT: We should refer to the companion (indicators) paper for more details.
% Also, I'm wondering what the point of the table (below) is?  Does it offer new
% info in addition to the paragraph?  If it's just a nice visual cue, then we
% can turn the above paragraph into a bullet point list, one bullet per signal,
% and cut the table. @DM/JB: Let me know what you think.
% DJM: agreed, table gone for now. I'll leave the bullet decision for later.

% \begin{table}
% \begin{tabular}{@{}p{1in}p{1in}p{1in}p{3in}@{}}
%  \toprule
%     Data source &   Signal name &   Abbreviation    &  
%         Description\cite{DelphiAPI}\\
%  \midrule
%    \attn{COVIDcast Survey via Facebook} &   Smoothed CLI-in-community & FBCLI &
%         Estimated percentage of people reporting illness in their local community,
%         including their household   \\
%     Change Healthcare   &   Smoothed adjusted oupatient CLI  &   CHCLI   &
%         Estimated percentage of outpatient doctor visits primarily about
%         COVID-related symptoms, with day-of-week effects removed  \\
%     Change Healthcare   &   Smoothed adjusted oupatient COVID  &   CHCOV &
%         Estimated percentage of outpatient doctor visits with confirmed
%         COVID-19, with day-of-week effects removed  \\
%     Google Symptoms     & Smoothed sum of anosmia and ageusia   &   GSSAA   &
%         The sum of Google search volume for anosmia and ageusia related searches,
%         in an arbitrary units that are normalized for overall search users\\
%  \bottomrule
% \end{tabular}
% \caption{Syndromic signals available in COVIDcast.\label{tab:signals-used}}
% \end{table}

% RJT: We need to explain WHY HRRs.  This is not familiar to basically anybody!
% It's an invention of academics who work in healthcare policy, from Dartmouth.
% We should cite them: https://www.dartmouthatlas.org/faq/.  Here's the
% rationale for us using them: 
% - They are groups of counties driven by the same hospital referral system.
% - They are somewhat homogenous population-wise.  Don't have a long tail of 
% many HRRs with small populations (like counties).  We should check this!  And
% could even the distribution vs the distribution with respect to counties (or
% some other comparison point). 
% - They would be more relevant to **hospitalization forecasting** given their
% definition, but our indicators should provide bigger boosts to case forecasts.
% (Hospitalizations | cases shouldn't have much dependence with indicators.
% Here I'm using this as in X | Y, i.e., X conditional on Y).
% - But to the extent that case projections inform hospitalization projects,
% this is still quite relevant & sensible to use HRRs. 
% @DM/JB: Please take a stab at this.
% DJM: stabbed.

As for geographic resolution, we consider prediction of COVID case rates (and
hotspots) aggregated at the level of an individual hospital referral region
(HRR). HRRs correspond to groups of counties within the same hospital referral
system. The Dartmouth Atlas of Healthcare Policy~\cite{DartmouthHRR}, defines
these 306 regions based on a number of characteristics. They are contiguous
regions containing at least one city where major cardiovascular or neurological
precedures are performed. Each HRR has a minimum population of about 120,000, and
the majority of hospital services for that population are performed by hospitals
within the region. Some HRRs are quite large (Los Angeles has
about 10 million), but generally HRRs are much more homogenous than the
$\sim$3100 counties 
would be but at much higher resolution than states. They would be more relevant for
hospitalization forecasting, given their definition. We have chosen to focus on
forecasting cases at this resolution because the indicators should provide a
bigger boost. Once we observe case information, there should not be much
dependence of future
hospitalizations on the indicators, but predicting cases should provide more timely
warnings of future hospitalization needs as opposed to waiting for the case
numbers.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/hrr_pop_histogram.pdf}
  \caption{Histogram of population size for HRR. Note that the $x$-axis is on
    the log scale.}
  \label{fig:hrr-histogram}
\end{figure}

>>>>

\title{Can Auxiliary Indicators Improve COVID-19 Forecasts and Hotspot
  Prediction?}

@@

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many available in Delphi's COVIDcast
API, we consider the following five auxiliary indicators: 
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Tracking and Impact Survey CLI-in-community (CTIS-CLI-in-community): 
  The estimated percentage of the population who knows someone who is sick,
  based on Delphi's surveys of Facebook users.
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
  COVID-19 related Google search volume for queries related to anosmia (loss of
  smell), ageusia (loss of taste), or both.  
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

For more details on the five indicators (including how these are precisely
computed from the underlying data streams) we refer to
\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html},
which documents all of the COVIDcast signals, as well our companion paper on 
COVIDcast API and database. For CTIS in particular, we refer to our companion 
paper on this survey. For the Google Search Trends data set, see
\cite{GoogleSymptoms}, and \cite{Klopfen:2020, Vaira:2020} for articles on the
relevance of anosmia and ageusia.

As for geographic resolution, we consider the prediction of COVID-19 case rates
and hotspots (to be precisely defined shortly) aggregated at the level of an
individual hospital referral region (HRR). HRRs correspond to groups of counties
in the United States within the same hospital referral system. The Dartmouth
Atlas of Healthcare Policy~\cite{DartmouthHRR}, defines these 306 regions based
on a number of characteristics. They are contiguous regions containing at least
one city where major procedures (cardiovascular or neurological) are
performed. Each HRR has a minimum population of about 120,000, and the majority
of hospital services for that population are performed by hospitals within the
region. Some HRRs are quite large (such as the one containing Los Angeles, which
has about 10 million people), but generally HRRs are much more homogenous in
size than the (approximately) 3200 counties are, and offer a nice middle ground
in between counties and states.

HRRs, by their definition, would be most relevant for hospitalization
forecasting.  We have chosen to focus on forecasting cases at this resolution 
because the indicators that we consider should intuitively be more useful in
latter task.  Predicting case rates and case hotspots at the HRR level is still
a reasonable goal, especially predicted case information could then be fed
downstream into hospitalization forecasting models.
